Supporting 10 common mistakes article

### 2025-09-01 example_10m.Rmd
shallow gene annotations

suppressPackageStartupMessages({
  library("getDEE2")
  library("DESeq2")
  library("kableExtra")
  library("clusterProfiler")
  library("eulerr")
  library("gplots")
  library("mitch")
  library("fgsea")
  library("RhpcBLASctl")
})

@@

>>>>

suppressPackageStartupMessages({
  library("getDEE2")
  library("DESeq2")
  library("kableExtra")
  library("clusterProfiler")
  library("eulerr")
  library("gplots")
  library("mitch")
  library("fgsea")
  library("RhpcBLASctl")
  library("msigdb")
  library("ExperimentHub")
})

@@

## Mistake 8 - using shallow gene annotations

```{r,m8a}

msdb <- gmtPathways("msigdb.v2025.1.Hs.symbols.gmt")
kegg <- msdb[grep("KEGG",names(msdb))]
kegg_med <- msdb[grep("KEGG_MED",names(msdb))]
kegg <- kegg[grep("KEGG_MED",names(kegg),invert=TRUE)] #separate KEGG LEG from KEGG MED
gobp <- msdb[grep("GOBP",names(msdb))]
reactome <- msdb[grep("REACTOME",names(msdb))]
gslist <- list("KEGG"=kegg,"KEGGM"=kegg_med,"Reactome"=reactome,"GOBP"=gobp,"MSigDB"=msdb)

# no. gene sets
d1 <- unlist( lapply(gslist,length) )

# total number of annotations
d2 <- unlist( lapply(gslist,function(i) { length(unlist(i)) } ) )

# median gene set size
d3 <- unlist(lapply(gslist,function(i) { median(unlist(lapply(i,length))) } ))

# number of genes with one or more functions
d4 <- unlist( lapply(gslist,function(i) { length(unique(unlist(i))) } ))

d5 <- unlist(lapply(gslist,function(i) { median(unlist(lapply(i,length))) } ))

data.frame(d1,d2,d3,d4) %>%
  kbl(caption="Gene set metrics") %>%
  kable_paper("hover", full_width = F)

kegg_up <- fora(pathways=kegg, genes=up, universe=bg, minSize = 5, maxSize = Inf)
kegg_dn <- fora(pathways=kegg, genes=dn, universe=bg, minSize = 5, maxSize = Inf)
nrow(subset(kegg_up,padj<0.05))
nrow(subset(kegg_dn,padj<0.05))
sum(nrow(subset(kegg_up,padj<0.05)), nrow(subset(kegg_dn,padj<0.05) ))

kegg_med_up <- fora(pathways=kegg_med, genes=up, universe=bg, minSize = 5, maxSize = Inf)
kegg_med_dn <- fora(pathways=kegg_med, genes=dn, universe=bg, minSize = 5, maxSize = Inf)
nrow(subset(kegg_med_up,padj<0.05))
nrow(subset(kegg_med_dn,padj<0.05))
sum(nrow(subset(kegg_med_up,padj<0.05)), nrow(subset(kegg_med_dn,padj<0.05) ))

reactome_up <- fora(pathways=reactome, genes=up, universe=bg, minSize = 5, maxSize = Inf)
reactome_dn <- fora(pathways=reactome, genes=dn, universe=bg, minSize = 5, maxSize = Inf)
nrow(subset(reactome_up,padj<0.05))
nrow(subset(reactome_dn,padj<0.05))
sum(nrow(subset(reactome_up,padj<0.05)), nrow(subset(reactome_dn,padj<0.05) ))

gobp_up <- fora(pathways=gobp, genes=up, universe=bg, minSize = 5, maxSize = Inf)
gobp_dn <- fora(pathways=gobp, genes=dn, universe=bg, minSize = 5, maxSize = Inf)
nrow(subset(gobp_up,padj<0.05))
nrow(subset(gobp_dn,padj<0.05))
sum(nrow(subset(gobp_up,padj<0.05)), nrow(subset(gobp_dn,padj<0.05) ))

msdb_up <- fora(pathways=msdb, genes=up, universe=bg, minSize = 5, maxSize = Inf)
msdb_dn <- fora(pathways=msdb, genes=dn, universe=bg, minSize = 5, maxSize = Inf)
nrow(subset(msdb_up,padj<0.05))
nrow(subset(msdb_dn,padj<0.05))
sum(nrow(subset(msdb_up,padj<0.05)), nrow(subset(msdb_dn,padj<0.05) ))
